---
title: "TRGN510_Final_Project"
author: "Mengyuan Zhang"
date: "10/25/2020"
output:
  html_document: default
  pdf_document: default
---

# Check current working directory
```{r}
getwd()
```

# Set working directory
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_young")
```

# Load the packages that will be used in the following steps 
```{r}
library(R.utils)
library(stringr)
library(dplyr)
library(Homo.sapiens)
library(gsubfn)
```



```{r}
options(stringsAsFactors = FALSE)
```

# Load and integrate the data of young group
Unzip the original data folder of young group
```{r}
#you may need the following code to unzip the original data
#R.utils::gunzip("gdc_download_20201027_215715.193634.tar.gz")
```

Untar the original data folder
```{r}
#you may need the following code to untar the original data
#utils::untar("gdc_download_20201027_215715.193634.tar", exdir = ".")
```

Create a new folder named "row_young_data" 
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project")
dir.create('raw_young_data')
```

Find all of files contained in the original data folder, and copy all of the files into the new folder we created last step.
```{r}
for (dirname in dir('raw_young')){
  file <- list.files(paste0(getwd(),'/raw_young/',dirname), pattern = '*.counts')
  file.copy(paste0(getwd(),'/raw_young/',dirname,'/',file), 'raw_young_data')
}
```

```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_young_data")
```

Because the original names of files are so complicated, I am going to give these files new names for this project to easily indentify them. 
Create list1 that containd 20 file names.
```{r}
list1 <- list.files("/Users/zmyyy/Desktop/510/final_project/raw_young_data", pattern = '*.gz')
list1
```

Create list2 that containd the 20 new names
```{r}
list2 <- str_c(c("young"), c(1:20), sep = "_")
list2
```

Create a data frame that included the original names and the new names, rename the names of the columns
```{r}
young_name <- as.data.frame(list(list1,list2))
colnames(young_name) <- c("file_names", "project_names")
```

Read the first file as a test to check the format
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_young_data")
test_young <- data.table::fread(young_name$file_names[1])
```

Create a data frame containing all data of the 20 files
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_young_data")
young_data <- data.frame(matrix(NA,nrow(test_young),nrow(young_name)))
for ( i in 1:nrow(young_name)) {
  print(i)
  young_data[,i] = data.table::fread(young_name$file_names[i])[,2]
}
colnames(young_data) <- young_name$project_names
```

Create a data frame containing all ensembl IDs
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_young_data")
ensembl_id <- data.table::fread(young_name$file_names[i])$V1
```

Combine the the data and ensembl IDs
```{r}
young_data <- cbind(ensembl_id = ensembl_id, young_data)
```

Check the last 6 rows data
```{r}
tail(young_data)
```

The last 5 rows data are different from the others, so delete the last 5 rows
```{r}
young_data <- young_data[1:(nrow(young_data)-5),]
```

Check if we successfully delete the last 5 rows
```{r}
tail(young_data)
```

# Load and integrate data of the old group
Do the same steps with the young group
```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_old")
#you may need the following code to unzip the original data
#R.utils::gunzip("gdc_download_20201027_220002.770079.tar.gz")
```


```{r}
#you may need the following code to untar the original data
#utils::untar("gdc_download_20201027_220002.770079.tar", exdir = ".")
```

```{r}
setwd("/Users/zmyyy/Desktop/510/final_project")
dir.create('raw_old_data')
```

```{r}
for (dirname in dir('raw_old')){
  file <- list.files(paste0(getwd(),'/raw_old/',dirname), pattern = '*.counts')
  file.copy(paste0(getwd(),'/raw_old/',dirname,'/',file), 'raw_old_data')
}
```

```{r}
list3 <- list.files("/Users/zmyyy/Desktop/510/final_project/raw_old_data",pattern = '*.counts')
list3
```

```{r}
list4 <- str_c(c("old"), c(1:20), sep = "_")
list4
```

```{r}
old_name <- as.data.frame(list(list3,list4))
colnames(old_name) <- c("file_names", "project_names")
```


```{r}
setwd("/Users/zmyyy/Desktop/510/final_project/raw_old_data")
test_old <- data.table::fread(old_name$file_names[1])
old_data <- data.frame(matrix(NA,nrow(test_old),nrow(old_name)))
for ( i in 1:nrow(old_name)) {
  print(i)
  old_data[,i] = data.table::fread(old_name$file_names[i])[,2]
}
colnames(old_data) <- old_name$project_names
```


```{r}
old_data <- cbind(ensembl_id = ensembl_id, old_data)
```

```{r}
tail(old_data)
old_data <- old_data[1:(nrow(old_data)-5),]
```

```{r}
tail(old_data)
```

# Combine the data of young group and old group into a usable format
```{r}
project_data <- merge(young_data, old_data, by = "ensembl_id")
```

Ensemnl_id in project_data contains ".", I am going to remove the "." in order to macth the ENSEMBL IDs with gene names
```{r}
ENSEMBL<- gsub("\\.\\d*", "", project_data$ensembl_id)
head(ENSEMBL)
```

Select gene names from Homo sapiens
```{r}
genes <- select(Homo.sapiens, keys=ENSEMBL, columns=c("SYMBOL", "TXCHROM"), 
                keytype="ENSEMBL")
```

The rows number do not match the row number of project_data, so delete the duplicated gene names
```{r}
genes <- genes[!duplicated(genes$ENSEMBL),]
```

Due to the ensembl_id in project_data do not macth the ENSEMBL IDs in genes, combine the ENSEMBL and project_data to make project_data_1 contain the same 
```{r}
project_data_1 <- cbind(ENSEMBL, project_data)
```

Combine the genes and project_data_1 to generat the data frame that will be used in followng steps
```{r}
project_data_1 <- merge(genes,project_data_1, by = "ENSEMBL")
head(project_data_1)
```

# Data pre-processing
```{r}
library(limma)
library(edgeR)
```

```{r}
group <- c(rep("young",20), rep("old",20))
```

```{r}
y <- DGEList(counts = project_data_1[5:44], group = group)
```


```{r}
cpm <- cpm(y)
```


```{r}
lcpm <- cpm(y, log=TRUE)
```


```{r}
L <- mean(y$samples$lib.size) * 1e-6
M <- median(y$samples$lib.size) * 1e-6
c(L, M)
summary(lcpm)

```

